Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization.

### 2023-02-10 manuscript.Rmd
add missing citations

The assumption of rationality can be connected to Bayesian reasoning and leads to our adoption of the framework.
Our rationale is as follows.
Researchers who decide to conduct a certain study are actually choosing a study to bet on.
They have to "place the bet" by conducting the study, therefore, invest resources and stand to gain epistemic value with some probability.
This conceptualization of choosing a study as a betting problem allows us to apply a "Dutch Book" argument [@christensenCleverBookiesCoherent1991].
This argument states that any better must follow the axioms of probability to avoid being "irrational," i.e., accepting bets that lead to sure losses.
Fully developing a Dutch book argument for this problem requires careful consideration of what kind of studies to include as possible bets, defining a conversion rate from the stakes to the reward, and modelling what liberties researchers have in what studies to conduct.
Without deliberating these concepts further, we find it persuasive that researchers should not violate the axioms of probability if they have some expectation about what they stand to gain with some likelihood from conducting a study.
The axioms of probability are sufficient to derive the Bayes formula, on which we will heavily rely for our further arguments.
The argument is not sufficient, however, to warrant conceptualizing the kind of epistemic value we reason about in terms of posterior probability; that remains a leap of faith.
However, the argument applies to any reward function that satisfies the "statistical relevancy condition" (XXX).
That is epistemic value is gained if evidence is observed, given that it is more likely to observe evidence when the theory holds than if the theory does not hold.

@@

The posterior probability $P(H|E)$ is of great relevance since it is often used directly or indirectly as a measure of corroboration of a hypothesis.
In the tradition of Carnap (XXX), in its direct use, it is called corroboration as firmness; in its relation to the a priori probability $P(H)$, it is called increase in firmness.
As noted before, we concentrate on posterior probability as a measure of epistemic value, since no measure shows universally better properties than others.
However, it is generally expected that any measure of corroboration increases monotonically with an increase in posterior probability $P(H|E)$.
We assume that any measure shows gain if, and only if, evidence is observed assuming $P(E|H) > P(E|\neg{}H)$.
This assumption is sometimes called statistical relevance condition (XXX).

@@

Having discussed that increasing the theoretical risk will increase the epistemic value, it is intuitive to task preregistration with maximizing theoretical risk, i.e., a confirmatory research agenda.
Indeed, limiting the type-I error rate is commonly stated as a goal of preregistration (XXX).
We argue that while such a conclusion is plausible, we must first consider at least two constraints that place an upper bound on the theoretical risk.

@@

To summarize, we showed that higher theoretical risk and lower uncertainty about theoretical risk lead to higher expected epistemic value across a variety of measures.
The former result, that increasing theoretical risk leads to higher expected epistemic value, reconstructs the appeal of confirmatory research.
However, theoretical risk is something researcher have only limited control over, e.g., theories are often vague and illdefined, resources are limited, and increasing theoretical risk usually decreases detectability (a special instance of this trade off is the well known tension between alpha error and power).
Before researchers may preregister a confirmatory study with high theoretical risk, they must ensure that they are able to conduct a confirmatory study.
Researchers must develop theories sufficiently and then have the resources to adequately balance detectability against theoretical risk.
A preregistration may at most help to invest some time into developing more specific, hence riskier, implications of a theory, but enabling confirmatory research is a task that requires action from the whole research community.
Incentive structures must be more willing to evaluate not the outcomes of a study but the rigor with which it was conducted.
Journal editors could encourage theoretical developments and be willing to accept registered reports.
Funding agencies should demand an explicit statement about theoretical risk in relation to detectability and must be willing to provide the necessary resources to reach adequate levels of both.

>>>>

The assumption of rationality can be connected to Bayesian reasoning and leads to our adoption of the framework.
Our rationale is as follows.
Researchers who decide to conduct a certain study are actually choosing a study to bet on.
They have to "place the bet" by conducting the study, therefore, invest resources and stand to gain epistemic value with some probability.
This conceptualization of choosing a study as a betting problem allows us to apply a "Dutch Book" argument [@christensenCleverBookiesCoherent1991].
This argument states that any better must follow the axioms of probability to avoid being "irrational," i.e., accepting bets that lead to sure losses.
Fully developing a Dutch book argument for this problem requires careful consideration of what kind of studies to include as possible bets, defining a conversion rate from the stakes to the reward, and modelling what liberties researchers have in what studies to conduct.
Without deliberating these concepts further, we find it persuasive that researchers should not violate the axioms of probability if they have some expectation about what they stand to gain with some likelihood from conducting a study.
The axioms of probability are sufficient to derive the Bayes formula, on which we will heavily rely for our further arguments.
The argument is not sufficient, however, to warrant conceptualizing the kind of epistemic value we reason about in terms of posterior probability; that remains a leap of faith.
However, the argument applies to any reward function that satisfies the "statistical relevancy condition" [@wesleyc.salmonStatisticalExplanation1970; @fetzerStatisticalExplanations1974].
That is epistemic value is gained if evidence is observed, given that it is more likely to observe evidence when the theory holds than if the theory does not hold.

@@

The posterior probability $P(H|E)$ is of great relevance since it is often used directly or indirectly as a measure of corroboration of a hypothesis.
In the tradition of Carnap, in its direct use, it is called corroboration as firmness; in its relation to the a priori probability $P(H)$, it is called increase in firmness [@carnapLogicalFoundationsProbability1950, preface to the 1962 edition].
As noted before, we concentrate on posterior probability as a measure of epistemic value, since no measure shows universally better properties than others.
However, it is generally expected that any measure of corroboration increases monotonically with an increase in posterior probability $P(H|E)$ and our argument applies to those measures as well.
We assume that any measure shows gain if, and only if, evidence is observed assuming $P(E|H) > P(E|\neg{}H)$.
This assumption is sometimes called statistical relevance condition [@wesleyc.salmonStatisticalExplanation1970; @fetzerStatisticalExplanations1974].

@@

Having discussed that increasing the theoretical risk will increase the epistemic value, it is intuitive to task preregistration with maximizing theoretical risk, i.e., a confirmatory research agenda.
Indeed, limiting the type-I error rate is commonly stated as a goal of preregistration [@nosekPreregistrationRevolution2018; @oberauerPreregistrationForkingPath2019; @rubinDoesPreregistrationImprove2020].
We argue that while such a conclusion is plausible, we must first consider at least two constraints that place an upper bound on the theoretical risk.

@@

To summarize, we showed that higher theoretical risk and lower uncertainty about theoretical risk lead to higher expected epistemic value across a variety of measures.
The former result, that increasing theoretical risk leads to higher expected epistemic value, reconstructs the appeal of confirmatory research.
However, theoretical risk is something researcher have only limited control over, e.g., theories are often vague and illdefined, resources are limited, and increasing theoretical risk usually decreases detectability (a special instance of this trade off is the well known tension between alpha error and power).
Before researchers may preregister a confirmatory study with high theoretical risk, they must ensure that they are able to conduct a confirmatory study.
Researchers must develop theories sufficiently and then have the resources to adequately balance detectability against theoretical risk.
A preregistration may at most help to invest some time into developing more specific, hence riskier, implications of a theory, but enabling confirmatory research is a task that requires action from the whole research community.
Incentive structures must be more willing to evaluate not the outcomes of a study but the rigor with which it was conducted [@caganSanFranciscoDeclaration2013; @schonbrodtResponsibleResearchAssessment2022].
Journal editors could encourage theoretical developments and be willing to accept registered reports [@friedLackTheoryBuilding2020; @friedTheoriesModelsWhat2020; @vanrooijTheoryDevelopmentRequires2020; @vanrooijTheoryTestHow2021].
Funding agencies should demand an explicit statement about theoretical risk in relation to detectability and must be willing to provide the necessary resources to reach adequate levels of both [@kooleRewardingReplicationsSure2012].
